### 2019-06-19 latex/afm-cs-main.tex
Final touches before sean and lucy see it.

\documentclass[twocolumn,oneside]{IEEEtran/IEEEtran}
% \documentclass[defaultstyle,11pt]{IEEETran}

\usepackage{showframe}
\usepackage{amssymb}		% to get all AMS symbols
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{graphicx}		% to insert figures
\usepackage{multirow}
\usepackage{cite}
\usepackage{listings}
\usepackage{booktabs}
\usepackage[percent]{overpic}
\usepackage{subcaption}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage[inkscapelatex=false, inkscapepath=svgsubpath]{svg}
\usepackage{url}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{placeins}
\usepackage[disable]{todonotes}

% Indent paragraphs inside enumerate
\usepackage{enumitem}
\setlist{  
  listparindent=\parindent,
  parsep=0pt,
}
\newcommand{\xyc}{\ensuremath{XY}\xspace}
\newcommand{\xc}{\ensuremath{X}\xspace}
\newcommand{\yc}{\ensuremath{Y}\xspace}
\newcommand{\zc}{\ensuremath{Z}\xspace}

\newcommand{\rzup}{\ensuremath{r_{z,\textrm{up}}}\xspace}
\newcommand{\rzs}{\ensuremath{r_{z,\textrm{s}}}\xspace}

\usepackage{catchfilebetweentags} % for data input
\newcommand{\scanD}[1]{\ExecuteMetaData[cs_data.txt]{#1}}
\newcommand{\baseline}[1]{\ExecuteMetaData[baseline_errors.txt]{#1}}

% *******************************************************************************
% This artificially increases the papersize and margins, so todo notes are legible.
% ***************** WARNING: comment this out for final version. ****************
%\usepackage[paperwidth=734.295pt, textwidth=516pt,
%            marginparsep=5pt, hoffset=10pt, marginparwidth=90pt,
%           textheight=696pt]{geometry}
% Use layouts to display the original dimensions, so we can get as close to that as possible with
% our fake page size.
%\usepackage{layouts}
%textwidth: \printinunitsof{in}\prntlen{\textwidth}
%linewidth: \printinunitsof{in}\prntlen{\linewidth}
% textwidth: \printinunitsof{in}\prntlen{\textheight}

% *******************************************************************************

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%       BEGIN DOCUMENT...         %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\title{Improving the Image Acquisition Rate of an Atomic Force
  Microscope Through Sub-sampling and Reconstruction}

\author{Roger A. Braker, Yufan Luo, Lucy Y. Pao and Sean B. Andersson
  \thanks{The authors are with the Dept. of Electrical, Computer, and
    Energy Engineering at the University of Colorado, 425 UCB,
    Boulder, CO 80309, United States. Phone: +1 (303) 492-2360. Fax:
    +1 (303) 492-2758. R. A. Braker (corresponding author
    roger.braker@colorado.edu) is a graduate student and L.Y. Pao
    (pao@colorado.edu) is the Richard \& Joy Dorf Professor.}
  \thanks{This work was supported in part by the US National Science
    Foundation (NSF Grant CMMI-1234980) and Agilent Technologies,
    Inc.} }

\maketitle
\begin{abstract}
  Undersampling-based approaches in Atomic Force Microscopy (AFM) aim
  to reduce the time to acquire an image by reducing the number of
  measurements needed while still maintaining image quality. In this
  paper, we describe a hardware implementation and demonstration of
  one such approach based on the use of collections of short scans
  known as $\mu$--paths. Using a commercial AFM, we acquire data on a
  grating sample. Reconstructions are made from these data using
  % three approaches: (1) inpainting, an interpolation technique that
  % diffuses information from sampled locations into unsampled
  % regions, (2)
  % basis-pursuit, a compressive sensing-based algorithm that seeks to
  % minimize a
  % measure of sparsity in the underlying image, and (3)
  a new variant of basis pursuit, Basis Pursuit with Vertical
  Variation (BPVV), that is designed to reduce artifacts arising from
  the sampling pattern. The quality of the resulting images is
  compared to images from standard raster scans of the same regions at
  comparable imaging rates using both the peak signal-to-noise ratio
  and the structured similarity index metric and a new metric we call
  the relative damage index. \todo{Re-write abstract, especially
    conclusions} These experiments demonstrate that at slow scan
  rates, the $\mu-$path scheme produces similar image quality with
  significantly less sampling (and thus less tip-sample interaction)
  and that at high rates, the undersampling scheme produces higher
  quality images than raster scanning.
\end{abstract}

\section{Introduction}\label{sec:introduction}
% textwidth: \printinunitsof{pt}\prntlen{\textwidth}

% linewidth: \printinunitsof{pt}\prntlen{\linewidth}

% paperwidth: \printinunitsof{pt}\prntlen{\paperwidth}

% marginparwidth: \printinunitsof{pt}\prntlen{\marginparwidth}

% marginsep: \printinunitsof{pt}\prntlen{\marginparsep}

% textheight: \printinunitsof{pt}\prntlen{\textheight}

The Atomic Force Microscope (AFM) is a powerful instrument capable of
imaging sample surface topography, material characteristics, and other
surface properties, with nanometer scale precision. AFMs acquire
information about the sample by monitoring the deflection of the
cantilever caused by the tip-sample interaction force. In general,
% AFMs acquire information about the sample through a variety of
% imaging modes, all of which rely on the deflection of the cantilever
% caused by the tip-sample interaction force. While the details of
% each of the modes are quite different, in general,
feedback control is used to hold the deflection signal constant and
information about the property of interest is inferred from the
applied control or the measured dynamics of the cantilever
\cite{Abramovitch:2007gt}. Because of its versatility, spatial
resolution, and ability to image in vacuum, air, and liquid, AFM is
widely used in a variety of disciplines, including physics, biology
and materials science
\cite{Dufrene:2017gm,Yang:2017im,Payton:2016hj,Altman:2015ic,Haase:2015eh}.

In addition to imaging static samples, AFM is used to study dynamics
in systems with nanometer-scale features
\cite{Yang:2017im,Shibata:2017da,Shibata:2015jd,Ando:2014ja}. The
image acquisition time of conventional AFMs, however, is typically on
the order of seconds to minutes, severely limiting the time-scales
that can be explored. Driven by the need for faster imaging, there
continue to be many active research efforts to overcome this
challenge. Since an AFM is fundamentally a mechanical microscope, many
of the approaches to high-speed AFM (HS-AFM) have focused on
modifications to the physical components. These have included the use
of small, fast cantilevers \cite{viani1999fast,
  braunsmann2010high,Adams:2016hg}, the development of faster
actuators \cite{Maroufi:2015gt,Yong:2012kd,Kenton:2012cm}, as well as
the application of advanced controllers
\cite{Rana:2018es,Yong:2015gr,butterworth2010adaptive,salapaka2002high}.
By combining different techniques, there are current generation,
high-end instruments that can image at rates on the order of 1-10
frames/sec \cite{Ando:2014ja}. However, the fastest rates are achieved
only over small scan ranges and there remain many systems of interest
whose dynamics are faster even than these instruments can achieve. In
addition, there is a large installed base of much slower AFMs that can
benefit from alternative approaches to improving imaging rate.

A complimentary class of HS-AFM approaches considers modifications to
the sampling scheme rather than the system mechanics. Alternative scan
trajectories like spirals, cycloids, and Lissajous figures, have been
used in place of the standard raster scan. These trajectories are
easier for the actuators to follow, allowing the tip to be moved more
quickly and leading to faster image acquisition on an otherwise
unmodified instrument
\cite{Bazaei:2017dm,Wu:2015dt,Rana:2014bj,Tuma:2012hv,Yong:2010gm}.
Such approaches are ultimately limited by the vertical bandwidth of
the instrument for regulating the deflection signal \cite{Teo:2016ev}.
Related to these are local scanning methods that use the measurements
in real-time to steer the tip to focus the scan on features of
interest, reducing imaging time by reducing the amount of sampling
needed \cite{Hartman:2018ja,Wen:2018fl,Zhang:2015cf,Huang:2014dw}.
While these have been shown to yield an order of magnitude or better
improvement in imaging rate, they are limited in the class of samples
that can be imaged.

An alternative group of non-raster scanning schemes, introduced in
\cite{song2011video,andersson2012non}, is based on the idea of
undersampling. As with local scanning, data acquisition time is
reduced by reducing the amount of measurements acquired. Unlike local
scanning methods, this approach still presents the full image. Taking
advantage of the redundancy in many natural signals of interest, the
final surface image can be recovered from the limited number of
measured pixels using a variety of different reconstruction methods
such as inpainting or schemes based on the theory of compressive
sensing (CS) \cite{chen2013enhancement,luo2015comparison}. In addition
to the reduced imaging time, undersampling schemes also reduce
tip-sample interactions, thereby reducing the likelihood of tip wear
or sample damage.

One simple way to create undersampling schemes is by modifying
existing full scan patterns including raster, spiral, and Lissajous
scanning. For example, subline sampling is generated by randomly
skipping some of the horizontal lines in raster scanning
\cite{han2018reconstruction,chen2013enhancement}. For spiral and
Lissajous patterns, the scan parameters (frequency and amplitude) can
be selected to ensure the trajectory only passes through a desired
fraction of the pixels in the final image. The scanning time for these
smooth undersampling patterns can be estimated based on the proportion
of the pixels in the trajectory. However, results in the literature of
CS make clear that \textit{randomness} in the sampling pattern is
essential for creating good reconstructions in the general case. The
smooth nature of these patterns reduces their randomness and thus
leads to less accurate reconstructions than a completely random
sampling of the pixels for the same sampling fraction.

In AFM, implementing a truly random pattern requires that the tip be
engaged with the surface to collect a measurement, lifted and moved to
the new location, and engaged again.
Because the re-engagement process
is typically slow, this can lead to excessively long image acquisition
times and negate the gains from undersampling \cite{andersson2012non}.
In \cite{maxwell2014compressed}, we introduced a discrete undersampling scheme,
called a $\mu$-path pattern, which consists of short randomly placed,
horizontal scans (see Fig.~\ref{fig:mu_mask} for one such pattern). It
is designed to balance the randomness needed (to ensure good
reconstruction) with continuous scanning (to reduce the number of tip
engagements).

Our previous work on the $\mu$-path pattern using theoretical
calculations, simulations, and a preliminary implementation,
demonstrates that the approach can achieve scanning time
reduction while maintaining faithful image reconstruction
\cite{maxwell2014compressed,Luo:2015tu, braker_hardware_2018}. 
The main contributions of this paper relative to that past work are:
\begin{itemize}
\item We introduce a new reconstruction algorithm which is designed to
minimize the artifacts arising from the structure of the $\mu$-path
scanning pattern. This algorithm is described and compared to existing
reconstruction methods through simulations in Sec. \ref{sec:reconstructionMethods}.
\item In our prior experimental prototype, \cite{braker_hardware_2018} we used very basic integral controllers for all three axes for both raster and $\mu$-path scans. Here, we design compensators for each axis using standard loop-shaping techniques (Sec.~\ref{sec:control}), which result in a substantial increase in closed-loop bandwidth (e.g., 50 Hz to 450~Hz for the $Z$-axis).
We show in Sec.~\ref{sec:ff_control} that one of the limitations of $\mu$-path scanning is the discontinuous $X$-axis reference (e.g., in moving between $\mu$-paths) which excites the cross-coupling modes between the $X$ and $Z$ axes, an affect which neither our prior experimental work nor the numerous simulation studies have accounted for. We mitigate the cross coupling with a feed-forward design.
\item We experimentally compare (Sec. \ref{sec:results}) $\mu$-path scanning to not only full density raster scans but also to course raster scans which are interpolated to a full image. This is in contrast to typical sub-line sampling which takes scan lines at random. The clear advantage to a course raster scan is that it requires no modification to the instrument, only to the post-processing.
\item The bulk of this work is based on experiment, not simulation, in contrast to prior work by ourselves \cite{Luo_nano_2015,maxwell_acc_2014} and others \cite{han_optimal_2018,jensen_reconstruction_2013,oxvig_structure_2017}. Our main focus here is on delineating the practical issues which must be addressed to effectively realize $\mu$-path scanning, which has only seen limited attention in the past.
\end{itemize}

\begin{figure}
  \centering
  \includegraphics[width=0.65\columnwidth]{figures-SBA/random_mask.pdf}
  \caption{An example of the horizontal $\mu$-path sampling pattern
    for a 256$\times$256 pixel image with 35 pixels in each short scan
    and a total of 20\% pixels scanned.}
  \label{fig:mu_mask}
\end{figure}
% =========================================================================
\section{Reconstruction methods}
\label{sec:reconstructionMethods}
Compressive sensing (CS) is a signal processing technique which aims
at signal reconstruction from a relatively small (sub-Nyquist limit)
number of measurements \cite{carmi2014compressive}. It takes advantage
of the approximate sparsity of real-world signals, that is, that many
coefficients of such signals are close to zero when represented in an
appropriate basis. CS methods seek the true image signal
$x\in\mathbb{R}^n$ from the observation equation,
\begin{equation}\label{op:observation}
  y = R x = RU\eta,
\end{equation}
\noindent where $y\in\mathbb{R}^m$ is the observation vector, $R$ is
an $m\times n$ matrix defining the measurements, $U$ is an $n\times n$
sparsity basis and $\eta$ is the sparse representation of $x$ in the
domain of $U$. In general, $m\ll n$. In an imaging application, where
the underlying signal is a matrix $X\in\mathbb{R}^{h\times s}$, we
take $x=\text{vec}(X)$, where the $\text{vec}(\cdot)$ operator stacks
the columns of a matrix.

\noindent
\begingroup \setlength{\tabcolsep}{1pt}
\begin{figure*}[ht!]
  \centering
  \begin{tabular}{cccccc}
    \textit{\small original image} & \textit{\small BPDN} & \textit{\small BPVV}
    & \textit{\small detail of original} & \textit{\small detail of BPDN} & \textit{\small detail of BPVV} \\
    \includesvg[width=0.16\textwidth]{figures/cs20ng_og.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_og_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv_zoom.svg}\\
    % 
    % ------------------------------
    % Row
    % 2
    % --------------------------------------------
    \includesvg[width=0.16\textwidth]{figures/dna_og.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bp.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_og_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bp_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bptv_zoom.svg}
  \end{tabular}
  \caption{Reconstruction comparison between BPDN and BPVV. (first
    column) Original raster-scanned image. (second column) BPDN
    reconstruction from random, 40 pixel long $\mu$-paths with 25\%
    sampling. (third column) BPVV reconstruction from the same
    sub-sampled data. (remaining columns) Corresponding details from
    the red boxes indicated in the raster image. The results show that
    BPVV reduces the artifacts arising from the horizontal scans of
    the $\mu$-path pattern that appear in BPDN reconstructions.}
  \label{fig:BPTV_demonstration}
\end{figure*}
\endgroup

In the AFM application, the probe can only measure a single pixel at a
time. Thus the rows of $R$ are a subset of the rows of an $n$ by $n$
identity matrix. Ideally, the sparsity basis and the measurement
matrix $R$ will have a low \textit{mutual coherence}, a measure of
how each of the rows of $R$ (the measurements) ``spreads
out'' in the domain of $U$ \cite{candes2007sparsity}. In the
following, we assume that $U^TU=I$, though in the simulation and
experimental results, we take $U^T$ as the Discrete Cosine Transform
(DCT). The DCT basis generally keeps a good balance between achieving
a low mutual coherence between $U$ and the $\mu$-path sensing matrices
$R$ and providing a high sparsity for typical AFM sample images.

Basis pursuit with denoising (BPDN) is one common realization of the
CS-based reconstruction problem, and is given by the optimization
\begin{equation}
  \min_{x} \left \| U^Tx \right \|_1 \quad
  \text{s.t.}\quad x\in Q_p = \{x:~\|Rx - y\|_2 < \sigma\} \label{op:bp}
\end{equation}
where $\sigma$ represents uncertainty in the measurements. BPDN
essentially searches for the sparsest signal from all the candidates
that match the measurements.

Although BPDN is effective in the general setting, using it for
reconstruction from horizontal $\mu$-path samples often yields
artifacts in the vertical direction (that is, the direction orthogonal
to the $\mu$-path scans) leading to strong discontinuities in the
image. These artifacts grow more prominent as the length of the
$\mu$-paths is increased \cite{maxwell2014compressed}, because these
longer paths lead to increased mutual coherence between the
measurement matrix $R$ and the sparsity basis $U$. However, longer
$\mu$-paths means fewer $\mu$-paths and thus shorter scan times.

To mitigate these affects, we propose a new variant of BPDN that we
term basis pursuit with vertical variation (BPVV). BPVV adds a
vertical total variation penalty in the spatial domain to the
optimization objective. That is, \eqref{op:bp} is modified to \todo{It
  has been my observation that some benefit is gained by including
  $D_h$, but weighting it less than $D_v$. Not sure if we should leave
  it in. If so, BPVV should be BPTV or something.}
\begin{align}
  \min_{x\in Q_p}~f(x) &= \min_{x\in Q_p} \left\|U^Tx\right\|_1
                         + \alpha\left\|D_vx\right\|_1
                         + \beta\left\|D_hx\right\|_1 \label{op:bp_tv}\\
                       &=\min_{x\in Q_p} \left\|W^Tx\right\|_1 \label{op:bp_tvW}
                       % \quad W
                       % = \begin{bmatrix}U&\alpha_hD^T_h&\alpha_vD^T_v \end{bmatrix}^T
                       % \nonumber
\end{align}
where
\begin{equation*}
  W = \begin{bmatrix}U^T&\alpha D_h&\beta D_v \end{bmatrix}
\end{equation*}
with ${D_v x=\text{vec}(\nabla_v X)}$ (resp.,
${D_v x=\text{vec}(\nabla_v X)}$) representing a discrete gradient of
a matrix in the vertical (resp., horizontal) direction (see, e.g.,
\cite[Sec. 6.1,]{becker_nesta_2011}).
% and
% \begin{align*}
%   (\nabla_v X)_{i,j} = 
%   \begin{cases}
%     X_{i+1,j} - X_{i,j}, & 1\leq i<h,~ 1\leq j\leq s,\\
%     0, & \phantom{:}i=h,~\phantom{:::::} 1\leq j\leq s.
%   \end{cases}\\
    %     (\nabla_h X)_{i,j} =
%   \begin{cases}
%     X_{i,j+1} - X_{i,j}, & 1\leq i\leq h,~1\leq j<s \\
%     0, & \phantom{}1\leq i\leq,~\phantom{:::} j=s
%   \end{cases}
% \end{align*}
Thus, in \eqref{op:bp_tv}, $\left \| D_v x \right \|_1$ (resp.,
$\left \| D_h x \right \|_1$) is the total variation (TV) of the
signal in the vertical (resp., horizontal) direction. The parameters
$\alpha$ and $\beta$ are weighting parameters. The description of BPVV
in \eqref{op:bp_tvW} can be interpreted as changing our assumption of
sparsity in the DCT basis to assuming sparsity in an overcomplete
dictionary \cite{candes_redundant_2011}. \todo{I believe this is
  correct: Sean/Yufan can you confirm or reject?} When written as
\eqref{op:bp_tvW}, the problem is in a form that may be solved with
NESTA~\cite{becker_nesta_2011}.

% For completeness, we give an overview here. To that end, we re-write
% the cost $f(x)$ in \eqref{op:bp_tvW} as
% \begin{equation*}
%   f(x) = \max_{u\in Q_d} \left\langle u, W^Tx \right\rangle,\quad Q_d = \{u:~||u||_{\infty} \leq 1\}.
% \end{equation*}

% One of the core features of the NESTA algorithm is to replace the
% non-smooth functional $f(x)$ with a smoothed approximation
% \begin{equation}
%   f_{\mu} = \max_{u\in Q_d} \left\langle u,
%     W^T x \right\rangle - \frac{\mu}{2}||u||_2^2. \label{eqn:fcn_prox}
% \end{equation}
% The gradient of \eqref{eqn:fcn_prox} is given by
% \begin{equation*}
%   \nabla f_{\mu}(x)[i] = Wu_{\mu}(x)
% \end{equation*}
% where
% \begin{equation*}
%   (u_{\mu}(x))[i] = 
% \begin{cases}
%   \mu^{-1}(W^Tx)[i], & |W^T x[i]| < \mu,\\
%   \text{sgn}(W^T x[i]), & \text{else}.
% \end{cases}
% \end{equation*}
% The optimization \eqref{op:bp_tv} may now be solved using
% Algorithm~\ref{ag:NESTA}.

% \begin{algorithm}
%   \SetKwInOut{Initialize}{Initialize} \SetKwInOut{Output}{Output}

%   \Initialize{$x \in Q_p, \mu=0.9||W^T x_0||_{\infty}$,
%   $L_{\mu} = \frac{1}{\mu}||W||^2_2$}
%     %   $x^{k+1} = x^{k}$\\
%   \While{$\mu < \mu_{\textrm{final}}$ } {
%   $k=0$\\
%   \While{$\Delta f_{\mu} > \delta$} {
%   $\alpha_k = \frac{1}{2}(k+1)$ \\
%   $\tau_k = \frac{2}{k+3}$ \\
%   $\displaystyle y_k = \text{arg } \min_{x\in Q_p}
%   \frac{L_{\mu}}{2}||x - x_k||_2^2 + \langle \nabla f(x_k),~ x\rangle$\\
%   $\displaystyle {z_k = \text{arg } \min_{x\in Q_p}
%   \frac{L_{\mu}}{2}||x-x_0||_2^2 +
%   \langle \sum_{i=0}^k \alpha_i \nabla f(x_i), ~x \rangle}$\\
%   $x_{k+1} = \tau_kz_k + (1-\tau_k)y_k$\\
%   $k \leftarrow k + 1$\\
% }
%   $x_0 = x_k$\\
%   $\mu = \gamma \mu$\\
%   $L_{\mu} = (1/\mu)||W||^2_2$ } \Output{Reconstruction $x_k$}
%   \caption{NESTA algorithm}
%   \label{ag:NESTA}
% \end{algorithm}
% \todo{seems like a waste of space to reprint the algo, since its
% really no different than the reference}

% In the termination condition for the inner loop,
% $\Delta f_{\mu} = |f_{\mu}(x_k) -
% \bar{f}_{\mu}(x_k)|/\bar{f}_{\mu}(x_k)$ where
% \begin{equation}
% \bar{f}_{\mu}(x_k) = \frac{1}{\min\{10, k\}} \sum_{\ell=1}^{\min\{10,k\}} f_{\mu}(x_{k-\ell})
% \end{equation}
% When $RR^T=I$, the sub-problems in lines 6 and 7 have a closed-form
% solution. Details can be found in~\cite{becker_nesta_2011}.

% % The sub-problems in lines 6 and 7 have the form
% % \begin{equation}
% %   p = \textrm{arg } \min_{x\in Q_p} \frac{L}{2}||x - r|| + \langle
% %   g,~x - r\rangle
% % \end{equation}
% % for some vectors $g$ and $r$ and, as shown
% % in~\cite{becker_nesta_2011}, have a closed-form solution given by
% % \begin{align}
% %   \lambda & = \textrm{max}(0, \sigma^{-1}\left| b -
% %     Rq\right| - L\\
% %   q &= r - L^{-1}g\\
% %   p &= \left(I - \frac{\lambda}{\lambda + L}
% %     R^TR\right)\left( \frac{\lambda}{L}R^Tb + q\right).
% % \end{align}

A simulation comparison between BPDN and BPVV is shown in
Fig.~\ref{fig:BPTV_demonstration}. The first column shows two raster
scans: the top is 512 by 512 pixel image of a CS20NG grating acquired
with an Agilent 5400 AFM and the bottom is a 256 by 256 pixel image of
DNA acquired with an Agilent 5500 AFM. For the CS20NG grating, we
sampled 15\% of the pixels with 50 pixel $\mu$-paths. We sampled 25\%
of the DNA image with 25 pixel long $\mu$-paths. The second columns
shows the full frame reconstructions using BPDN and the third columns
shows the result using BPVV. The fourth column zooms in on the region
indicated by the red box in column one. Similarly, the fifth and six
columns show the detail of the reconstructions. The reconstruction
results show the artifacts apparent in the BPDN reconstruction are
largely mitigated using BPVV. Thus, in the remainder of the paper, we
focus exclusively on BPVV. As for reconstruction time, it took about
1.5 seconds to reconstruct the DNA image and about 7 seconds for the
CS20NG.

% =========================================================================
\section{Experimental setup} \label{sec:experimentalSetup}
% =========================================================================
\begin{figure}[ht!]
  \includegraphics[width=1\columnwidth]{figures/exp_setup.pdf}
  \caption{Schematic diagram of the expermental AFM setup.}
  \label{fig:exp_setup}
\end{figure}

Our experimental setup, illustrated in Fig.~\ref{fig:exp_setup},
consists of an Agilent 5400 AFM retrofitted with an nPoint NPXY100A
piezoelectric stage, a cRIO-9082 embedded controller (National
Instruments), and a standard desktop computer.

Through a breakout box, the Agilent 5400 provides access to the
$Z$-axis deflection signal. When the Agilent software is set to
open-loop mode, a $\pm10$~volt input on the standard control box
allows control of the $Z$-axis of the N9524a piezo scanner, which has
a total range of 7 $\mu$m. Initial (course) engagement of the tip to
the sample is performed using the PicoView software before control is
handed over to the custom controllers.

All control logic is programmed using LabVIEW 2019 and compiled to a
Xilinx Spartan-6 LX150 Field Programmable Gate Array (FPGA) inside the
cRIO-9082. The cRIO includes a 16-Bit, 100 kHz NI-9215
analog-to-digital input module and a 16-Bit, 100 kHz NI-9263
digital-to-analog output module. All control loops are implemented
using a 25 kHz sampling rate.

% =========================================================================
\section{Implementation}\label{sec:implementation}
% =========================================================================

Implementing the $\mu$-path scheme involves operating the AFM in
several distinct stages. In the $XY$-direction, the system transitions
from tracking a step command in the transition to a new
measurement location to tracking a short ramp during the actual $\mu$-path scan.
In the $Z$-direction, the system must
transition between tip descent, surface scanning, and tip retraction.
Two cycles of this sequential process are illustrated by the time
series in Fig.~\ref{fig:cs_cycle}. The following
subsections describe in more detail several features of the CS
scanning process.

Transition between, and operation in, these different stages is
implemented as a simple state machine, summarized in
Table~\ref{tab:cs_tasks}. The main job of the state machine is to
adjust the reference signals to accomplish each task.
In the following, $r_X$ and $r_Y$ and $r_Z$ refer to
the references for the $X$ and $Y$ axes while $r_{Z,s}$ and $\rzup$
refer to the $Z$-axis setpoints during scanning and retraction,
respectively.

\begin{table}[h!]
  \centering
  \caption{Summary of the tasks for $\mu$-path scanning.}
  \begin{tabular}{lccc}
    state & $r_X$,$r_Y$ & $r_Z$ & next-state\\
    \toprule
    (1) $XY$-move & setpoint & $r_Z = \rzup$ & (2)\\
    (2) tip-engage & setpoint & $r_Z = \rzs$ & (3)\\
    (3) pre-scan & ramp & $r_Z = \rzs$ & (4)\\
    (4) $\mu$-path scan & ramp & $r_Z = \rzs$ & (5)\\
    (5) tip up & setpoint & $r_Z = \rzup$ & (1)\\
  \end{tabular}
  \label{tab:cs_tasks}
\end{table}

\subsection{Cantilever does not need to fully disengage.}
\todo{Not clear to me yet that friction is useful if I scan parallel
  to cantilever. Also not clear to me that the normal force and shear
  force are orthogonal concepts. Maybe the macroscopic analogy breaks
  down, but thinking of a knife pushed into a surface and then
  dragged...}
In our initial implementation \cite{braker_hardware_2018}, we insisted that the during
tip-retraction (state-5), the cantilever tip should fully break
contact with the surface. Here, we impose no such requirement. Rather,
we change the $Z$-setpoint during tip-retraction to only pull away far
enough that the deflection signal stays low (i.e., below the scanning
setpoint), even while we run across the surface. We choose \rzup so
that the interaction between the tip and sample becomes purely
adhesive. In general, stable imaging in is not possible at \rzup and
complete dis-engagement occasionally occurs. This can be seen in bottom pane of
Fig.~\ref{fig:cs_cycle} during the second $XY$-move (black portion) where the tip detaches partway through the move.
%Fig.~\ref{fig:afm_bd_dinv} which selects either the scanning setpoint $r_{Z,s}$ or the retraction setpoint $r_{Z,\textrm{up}}$ based the current state.
When dis-engagement does occur, the $Z$-axis control loop implements an anti-windup scheme to prevent the control from saturating.

Only moving the tip into the adhesive region during the $XY$ move saves considerable time
compared to completely detaching the tip between each $\mu$-path. By eliminating the normal force into the sample, we are able to take dozens of images with a single probe without a noticeable degradation in image quality. However, completely characterizing the damage done to the tip and sample with this scheme compared complete detachment (which can also damage the probe \cite{meyer_atomic_1992}), remains an outstanding goal.

\begin{figure}[t!]
  \centering \includesvg[width=1\columnwidth]{figures/cs_cycle.svg}
  \caption{Several CS cycles. Each state is indicated by color.}
  \label{fig:cs_cycle}
\end{figure}

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=1\textwidth]{figures/AFM_loop_xyz.pdf}
  \caption{Block diagram of the $Z$-axis control loop. The anti-windup
    is active for decision index $i\in\{0,1,5\}$, which correspond to
    states such that $r_Z=\rzup$.}
  \label{fig:afm_bd_dinv}
\end{figure*}
\subsection{The pre-scan}
After the tip-descent, we begin scanning in a ``pre-scan'' phase
(state (3)), which is the turquoise portion of the trajectories in Fig.~\ref{fig:cs_cycle}.
There are two reasons for this. First, we have observed that there is less noise in the deflection signal while the $XY$-stage is in motion compared to sitting still. More importantly, it lets us bring the $X$-axis to its scanning velocity at the same time the transients from the $Z$-axis die out. Without this, we must extend the $\mu$-path by a distance equal to the steady-state error following a ramp. For an arbitrary scan speed, this means the $\mu$-path scan would need an extra 
\begin{equation}
  N = \left\lceil\lim_{z\rightarrow 1} \frac{1-H_x}{z-1}\right\rceil
\end{equation}
time steps, where $H_x$ is closed-loop transfer function for the $x$-axis.
For our system $N=111$. By initiating the scan while the $Z$-axis settles, we eliminate much of
this overhead.

\subsection{Control}\label{sec:control}
% \begin{figure}
%   \centering
%   \includesvg[width=1\columnwidth]{figures/z_control_design.svg}
%   \caption{FRFs of the $Z$-axis. The blue curve is the open loop
%   response from $u_Z$ to deflection. By inverting the bending mode
%   at 215~Hz, we achieve the black curve in closed-loop.}
%   \label{fig:z_control}
% \end{figure}
The control structure for each axis, shown in the block diagram in
Fig.~\ref{fig:afm_bd_dinv}, is similar and essentially consists of a
set of filters in series with a PI controller. The open-loop frequency
response for the $X$ and $Y$ axes is shown as the black curves in Fig.~\ref{fig:mimo_frf_uxuy} and for the $Z$-axis in Fig.~\ref{fig:z_control}. The first resonance of each axis is the bending mode, which
appears at 215~Hz for the $Z$ axis and about 350~Hz for the $X$ and
$Y$ axes.

The filter portion of the $Z$-axis controller $D_Z$ is second-order
and simply inverts the complex-conjugate pole-zero pair at 215~Hz.
This allows the gain of the PI controller to be increased
substantially. The closed-loop achieves a bandwidth of about 450~Hz,
as shown by the red curve in Fig.~\ref{fig:z_control}. Without this
inversion, the bandwidth is limited to about 50~Hz.

We take a similar approach for the $X$ and $Y$ axes. The filters $D_x$
and $D_y$ invert the bending modes at 350~Hz and add notch filters at
the main resonant peaks near 650~Hz. The resulting closed-loop
transfer functions are shown as the red curves in
Fig.~\ref{fig:mimo_frf_uxuy}.

In general, the $X$-axis positioning requirements are more stringent
for both CS and raster scanning. Due to the step inputs the $X$-axis must track, we have found that
compensating hysteresis and creep to be particular helpful during $\mu$-path scanning. Thus,
the $X$-axis compensator also includes an inverse creep and hysteresis
model. The hysteresis model is a modified Prandtl-Ishlinki type model
and is based on the work of \cite{kuhnen_modeling_2003}, while the
drift compensation models the piezo creep as a simple second order
transfer function with real poles and zeros. In
Fig.~\ref{fig:afm_bd_dinv}, the hysteresis inversion is indicated by
$\mathcal{H}^{-1}[\cdot]$. We have found that the inclusion of these
extra compensators limits the overshoot of the $X$-axis during larger
moves between different $\mu$-paths (e.g., without them, the $X$-axis trajectory in Fig.~\ref{fig:with_ff} would show noticeable overshoot). More details on fitting the creep
and hysteresis models can be found in our other work \cite{braker_afmmpc_2019}.

\begin{figure*}
  \centering
  \begin{subfigure}[b]{.65\textwidth}
    \includesvg[width=1\textwidth]{figures/MIMO_CL_uxuy.svg}
    \caption{}\label{fig:mimo_frf_uxuy}
  \end{subfigure}\hfill
  \begin{subfigure}[b]{.34\textwidth}
    \includesvg[width=1\textwidth]{figures/with_wo_ff.svg}
    \caption{}\label{fig:with_ff}
    \vspace{2ex}
    \includesvg[width=1\textwidth]{figures/z_control_design.svg}
    \caption{}\label{fig:z_control}
  \end{subfigure}
  \caption{(a) MIMO closed and open-loop FRFs, showing the dramatic
    affect on $G_{Z,u_X}$ and $G_{Z,u_Y}$ when the feedforward
    compensators are used for the $X$ and $Y$ axes. (b) The vertical control $u_Z$ (top) after a large move in the $X$-direction (bottom) without (left column) and with (right column) the feedforward compensation. (c) Frequency response of the $Z$-axis in open-loop (black) and closed-loop (red). }
\end{figure*}

% We note that the $z$-axis control described here operates the AFM in
% constant force mode during imaging. It is straightforward to perform
% scanning under different AFM modes, such as tapping mode, simply by
% adjusting the signals and control involved in the vertical loop.

\subsection{Feedforward Control}\label{sec:ff_control}
Unfortunately, as the bottom row of plots in Fig.~\ref{fig:mimo_frf_uxuy}
makes clear, there is a strong cross-coupling from the horizontal
plane to the $Z$ axis (the coupling from $u_Z$ to the $X$ and $Y$ axes
is very weak, and so is not shown). Our SISO control scheme does
little to attenuate the resonances of $G_{Z,u_X}$ and $G_{Z,u_Y}$. In
particular, there are modes in both these transfer functions at 215
and 505 Hz that do not appear in a SISO model of $G_{X,u_X}$ or
$G_{Y,u_Y}$.

For raster scans, the frequency content of the $X$-axis reference
(which is a triangle wave) decays like $1/n^2$ with a gain
proportional to the raster frequency. In contrast, the $X$-axis
reference for a $\mu$-path scan is composed of a sequence of steps and
ramps. The inclusion of the (discontinuous) steps means we should
expect the $\mu$-path $X$ reference to decay like $1/n$. Thus, while
the cross-coupling can induce $Z$-axis vibrations for faster raster
scans, it is especially problematic for $\mu$-path scanning. This is
illustrated in the left column of Fig.~\ref{fig:with_ff}, which shows
a CS cycle during which the $X$-axis makes a 5 micron move, which
induces excessively large vibrations in the vertical control.

In principle, one could mitigate these issues with a full MIMO control
design. In this work, we opt for a much simpler feedforward scheme.
For CS scanning, we insert feed-forward filters ($F_X$ and $F_Y$ in
Fig.~\ref{fig:afm_bd_dinv}) with three notches: one at 215~Hz, one at
350~Hz and one at 505~Hz. Both filters are shown in Fig.~\ref{fig:mimo_frf_uxuy}
as the dashed-pink curves and the resulting overall FRF is shown as the
dotted-blue curves, which show a dramatic reduction in the cross
coupling. A similar affect could be achieve by using a simple low-pass
filter. However, to achieve the same 40~dB of attenuation of the mode
at 215 Hz, would require, e.g., that a two-pole low-pass filter have
its cut-off frequency at about 20~Hz. Thus, while our feedforward
scheme slows the $X$-axis down, we achieve better bandwidth than we
would with a simple low-pass filter.

For raster scanning, we set $F_X=F_Y=1$ and instead take advantage of
the periodicity of the $X$-axis reference, which is a triangle wave.
We modify the triangle wave to a truncated Fourier series, where each
Fourier coefficient is scaled by the inverse of the closed-loop FRF at
the corresponding frequency. See, e.g., \cite{clayton_review_2009} for
more details on this strategy. We truncate this series such that the
highest frequency is smaller than 200 Hz, which means that we should
never excite the problematic $G_{Z,u_X}$ modes. This scheme implies
that the effective bandwidth of the $X$-axis controller is higher
while raster scanning than while CS-scanning. On the other hand, it
also implies that the for faster scan rates, fewer Fourier
coefficients are used, and so the approximation to a triangle wave is
worse.

\section{Metrics}\label{sec:rdi}
In section~\ref{sec:results} we use three metrics to asses our
results. The first two seek to quantify image quality of two images.
Define a master image as $X$ and a reconstruction or corrupted version
as $Y$, with each having $s\times h=m$ pixels. Let $x=\textrm{vec}(X)$
and $y=\textrm{vec}(Y)$. Let $L$ be the dynamic range of the master
image $x$. Then the peak signal to noise ratio (PSNR) is given by
\begin{equation*}
  \text{PSNR}(x,y) = 10\log_{10}\frac{L^2}
  {\sqrt{\frac{1}{p^2} \sum_{i=1}^{m}( x_{i} - y_{i})^2 }}.
\end{equation*}

The goal of the SSIM is to compare two image's structure,
luminescence, and contrast and is built up from the means ($\mu_x$ and
$\mu_y$), standard deviations ($\sigma_x$ and $\sigma_y$), and
covariance ($\sigma_{xy}$) of the image vectors $x$ and $y$. The
variation of the SSIM used in this paper is defined as
\begin{equation*}
  \text{SSIM}(x,y) = \frac{(2\mu_x\mu_y + C_1)(2\sigma_{xy}+C_2)}
  {(\mu_x^2 + \mu_y^2 + C_1)(\sigma_x^2 + \sigma_y^2 + C_2)}
\end{equation*}
where the constants $C_1$ and $C_2$ are regularizing constants to
prevent singularity if, e.g., $\mu_x=\mu_y=0$. We use the default
values suggested in \cite{wang_image_2004} of $C_1=(0.01L)^2$ and
${C_2=(0.03L)^2}$, where again, $L$ is the dynamic range. In general
two images which are identical will yield an infinite PSNR and an SSIM
of 1.

\subsection{Limitations of SSIM and PSNR}\label{sec:limits}
The SSIM and PSNR have become reasonably popular in studies comparing
simulated AFM image reconstructions \cite{oxvig_structure_2017,
  han_optimal_2018, Luo_nano_2015}. However, it remains somewhat of an
open question how to best compare experimental AFM images. While we
use the SSIM and PSNR in Sec.~\ref{sec:results}, we believe those
numbers should be interpreted with some caution. In our experimental
setup, actuation in the $XY$-plane happens via the nPoint piezo stage
and the $XY$-direction of the original piezo scanner is uncontrolled.
This leads to substantial drift between images. To illustrate this, we
took a sequence of 6 raster scans of a sample grating, each at a 1~Hz
scan rate over a 5 micron by 5 micron area. The left image of
Fig.~\ref{fig:baseline_errors} shows the error between the first and
 sixth image. The PSNR and SSIM between
the first and second images are \baseline{one2twoNoAlgnPsnr} and
\baseline{one2twoNoAlgnSsim}; between the first and sixth are
\baseline{one2sixNoAlgnPsnr} and \baseline{one2sixNoAlgnSsim},
respectively, even though on their own each image appears to have good quality.

To mitigate this affect, we use an image alignment technique based on
the 2D cross-correlation between each image. Essentially, this takes a
sub-slice of the master (first) image (here, a 25 pixel inset on each edge), and finds
the offset in the other image that produces the smallest error. The
result of this procedure are shown in the right image of
Fig.~\ref{fig:baseline_errors}.
Although the situation is improved compared to the un-aligned images,
there is still noticeable stretching (the upper and lower portions appear considerably worse than the middle). Notably however, with the alignment technique, the SSIM is
\baseline{one2twoAlgnSsim} for both the second and sixth image images and
the PSNR is increases to \baseline{one2twoAlgnPsnr} and \baseline{one2sixAlgnPsnr}
respectively.

\begin{figure}[t!]
  \centering
  % \includesvg[width=0.7\columnwidth]{figures/baseline_errors_aligned_1Hz.svg}
  \includesvg[width=0.7\columnwidth]{figures/baseline_errors_1and6.svg}
  \caption{Errors between the first and sixth raster scans in a sequence of six 1.0~Hz
    raster scans. (left) Errors without alignment (right) The same
    data but aligned via cross correlation.}
  \label{fig:baseline_errors}
\end{figure}

\subsection{A new metric}
Th PSNR and SSIM metrics only assess the similarity of two images.
However, in AFM imaging, pure image quality is not the only concern,
particularly for delicate samples. What is needed is a figure of
merit for how much damage was done to specimen during the imaging
process. While damage to the specimen is not really a concern while imaging a hard
calibration grating, it plays a prominent role for biological
samples~\cite{ando_highspeed_2008}, though damage to the probe itself is still a concern.
In general, the faster the scan rate,
the more difficulty the $Z$-axis control
will have following the sample surface,
which results in a larger $Z_d$ over more time. 
The actual damage done by a given $Z_d$ will
depend on the spring constant of the cantilever and the softness of
the specimen. Nonetheless, we propose a metric, which we call the Relative Damage Index (RDI)
which computes the power in the deflection signal for a given scan. That is,

\begin{align}
  \text{RDI} &= \frac{1}{T_s}\sum_{k=0}^N \frac{1}{k} \left(Z_{d,k} - r_{Z,s}\right)^2 \label{eqn:RDI}\\
\end{align}
The idea behind the RDI is that increased scan rates lead to larger
deflection signals (and thereby increased force into the specimen) at
each surface feature, which the RDI accumulates.

\section{Experimental Results}\label{sec:results}
We compared $\mu$-path scanning to two scenarios. First, we compare it
to full resolution raster scans at several scan rates in Sec.~\ref{sec:cp_full}.
Second, we compare to course raster scans in Sec.~\ref{sec:cp_subline}.

\subsection{Comparison to full raster}\label{sec:cp_full}
We took scans of a Ted Pella CS-20NG sample grating over an area with
holes on a 500 nm pitch. The holes are 20 nm deep. We scanned 5 micron
by 5 micron images using standard raster scans with 512 lines. For the
raster scans, we took scans at 1.0~Hz, 2.5~Hz, 5.0~Hz, 8.0~Hz and
10~Hz. We also took $\mu$-path scans with densities at 12.5\%,
and 25\% and reconstructed 512 by 512 pixel images. The CS scans were
taken with scan velocities equivalent to raster scan rates of 1.0~Hz,
2.0~Hz, 5.0~Hz and 8.0~Hz. For all CS scans, the $\mu$-paths are
$\scanD{muLenNM}$ nm long, which translates to $\scanD{muLenPix}$
pixels.

The free value of the deflections signal was approximately -0.6 volts
and we set the scanning reference to ${\rzs=\scanD{spScan}}$~volts (for both raster and CS)
and the withdraw reference to ${\rzup=\scanD{spMv}}$ volts. We set the
$XY$-settle boundary to $\pm \scanD{xyBoundaryMic}$ microns and
required $\scanD{xySetSamps}$ samples within this boundary before
moving to the $Z$-engage state. The pre-scan length was
$\scanD{preScanSamples}$ time steps. The cantilever was a Budget
Sensors ContAI-G, with a nominal length of 450 microns and spring
constant of $k=0.2$~N/m.

% \SSData{G2}$& $g_6 = \SSData{G6}
The CS scans were reconstructed with BPVV. We set
${\alpha_v=\scanD{alpv}}$ and ${\alpha_h=\scanD{alph}}$. 
For the raster scans, we discard data from the re-trace and divided
the remaining data for each scan line into 512 bins based on the \xc
sensor measurements. We then average the data in each bin to obtain
the value of one pixel. To remove the affects of piezo creep and
sample tilt, we then de-trend each individual line. De-trending each
line causes the flat portion between the holes in the horizontal
direction to appear at a different height than the flat portion
between holes in the vertical direction. To remove this effect, we
select a column of pixels on the left and right side of each image
that does not cross any holes (i.e., a vertical line through the flat
area), register each scan line to a common height along these columns.
This step eases comparison with the CS scans and was inspired by a
feature in the AFM image processing toolbox \texttt{SPIW} \cite{spiw}.
The resulting images are shown in Fig.~\ref{fig:resultsF1_images}.
The color maps represent a range of $\pm$20 nm. The
rows of pixels indicated by the red lines are shown as cross sections
Fig.~\ref{fig:pixel_rows}.

\begin{figure*}
  \centering
  \includesvg[scale=.98]{figures/cs_raster_images_6-5-2019.svg}
  \caption{Raster and compressed sensing images of a 5 micron square
    area of the CS-20NG grating. All images are 512$\times$512 pixels:
    the raster images marked as 64 and 128 lines have been
    interpolated. The rows of pixels indicated by the red line are
    shown in Fig.~\ref{fig:pixel_rows}.}
  \label{fig:resultsF1_images}
\end{figure*}

% \begin{table}[t!]
%   \centering
%   \caption{Performance metrics for raster and CS scans.}
%   \begin{tabular}{cccccc}
%     \input{tables/cs_raster_table_6-5-2019_muInf_dct2.tex}
%   \end{tabular}
%   \label{tab:rast_vs_cs_v1}
% \end{table}

\begin{table}[t!]
  \centering
  \caption{Breakdown of state times for the CS scans shown in
    Fig.~\ref{fig:resultsF1_images}. All times are in seconds.}
  \label{tab:final_state_times}
  \begin{tabular}{ccccccc}
    \input{tables/cs_state_times_table_6-5-2019_muInf_dct2.tex}
  \end{tabular}
\end{table}

\begin{figure}
  \includesvg[width=1\columnwidth]{figures/cs_raster_pixel_rows_6-5-2019.svg}
  \caption{Rows of pixels, as indicated by the red lines in
    Fig.~\ref{fig:resultsF1_images}. For clarity, not all images are
    included.}
  \label{fig:pixel_rows}
\end{figure}
For each raster scan and each CS scan, we computed the SSIM and PSNR
 using the 1~Hz raster scan as the master and
also computed the RDI metric. These are plotted against total acquisition time
in Fig.~\ref{fig:time_ssim} and ~\ref{fig:time_psnr}.
Unfortunately, the SSIM and PSNR metrics are not particularly
illuminating. There are several inconsistencies. For example, it makes no sense
the 1~Hz CS scan at 25\% has a \emph{worse} PSNR and SSIM than 5~Hz version;
nor does it make sense that the 8~Hz full raster scan has a worse SSIM and PSNR than
the 10~Hz full raster scan.

\begin{figure*}[t!]
  \centering
  \begin{subfigure}{0.329\linewidth}
    \centering
    \includesvg[width=1\columnwidth]{figures/cs_rast_time_vs_ssim.svg}
    \caption{}
    \label{fig:time_ssim}
  \end{subfigure}
  \begin{subfigure}{0.329\linewidth}
    \centering
    \includesvg[width=1\columnwidth]{figures/cs_rast_time_vs_psnr.svg}
    \caption{}
    \label{fig:time_psnr}
  \end{subfigure}
  \begin{subfigure}{0.329\linewidth}
    \centering
    \includesvg[width=1\columnwidth]{figures/cs_rast_damage.svg}
    \caption{}
    \label{fig:time_damage}
  \end{subfigure}
  \caption{Metrics for raster scan at 512, 128 and 64 lines and
    $\mu$-path scans at 12.5\% and 25\% sampling. The scan rate is
    indicated by color while the scan style is indicated by marker
    style. (a) SSIM as a function of total acquisition time. (b) PSNR
    as a function of total acquisition time. (c) RDI as a function of
    total acquisition time.}
  \label{fig:time_vs_metrics}
\end{figure*}

Things are clearer if we plot the RDI for each scan against total acquisition time, which is
shown in Fig.~\ref{fig:time_damage} and 
this is where CS seems to shine.
As one would expect, as imaging time decreases, the RDI for each
scan type increases. 
From Fig.~\ref{fig:time_damage}, we see that for a
fixed RDI, CS has a faster imaging acquisition rate. The amount of
improvement is best if we need a very low RDI: for example, if we want
to hold the RDI below 2, then we have to take the 1~Hz raster scan
(512 seconds) while we can achieve the same RDI via CS in about 42
seconds with the 12.5\% sampling, which is a speed improvement of over
10. The higher of an RDI we are willing to tolerate, the advantage of
CS over raster narrows. For example, the 8~Hz raster scan and 8~Hz,
12.5\% CS scan have comparable RDIs, but now CS is only about 5 times
faster. This is a result of the constant overhead imposed by the
engage/disengage and $XY$-move states in CS which remains constant for
different scan rates.

\subsection{Comparison to Sub-line Sampling} \label{sec:cp_subline}
 Many of sub-line sampling papers assume
that the lines to sample are selected randomly \cite{han_optimal_2018,
  Luo_nano_2015}. An even simpler alternative to
$\mu$-path sampling is to simply 
take a standard raster scan with fewer lines. This is the technique we
consider in this subsection, by scanning images with 
 128 and 64 lines. For each of these scenarios, we first divide the data
for each line into 512 bins. This gives an image that is 128 (resp.,
64) pixels by 512 pixels. We then linearly interpolate along each
column to produce a 512 by 512 pixel image. Again, we take these
courser raster scans at 1.0, 2.5, 5.0, 8.0 and 10.0 Hz. The resulting
images are shown in Fig.~\ref{fig:results_subline_images}.

Both visually and from the metrics show in
From both From the metrics in Figs.Figs.~\ref{fig:time_ssim} and \ref{fig:time_psnr}, the 64 line scan
performs the worst of the all the sub-sampling methods and is slower than 12.5\% CS for slow scan rates. However, the 128 line scans appear to have as good or better image quality than either CS scans.

Note the defect on the hole located in the second column, sixth row
from the top (of the grating holes). The CS-scans do a poor job detecting this,
while both course raster scans do so. However, the other artifacts in 64 line
scan obscure the defect.

Fig.~\ref{fig:improve_512} compares the amount of speed improvement of the different sub-sampling scheme over a full raster scans with 512
lines as a function of scan rate. As far as pure speed goes, the CS
scan at 12.5\% is the fastest for a rate of 1~Hz, though the 64 line
raster scan is close. At the same time, the image quality of the 1~Hz,
12.5\% CS scan is better than the the 1~Hz 64 line raster scan.
However, this advantage quickly evaporates for faster scan rates.
Indeed, for rates beyond about 5~Hz, the raster scan with 128 lines is
not only faster than the 12.5\% CS scan, it also has (arguably) better
image quality.

\begin{figure*}
  \centering
  \includesvg[scale=.98]{figures/cs_raster_images_subl_6-5-2019.svg}
  \caption{Raster scans at 128 (top) and 64 (bottom) which have been
    interpolated along each column to 512 lines.}
  \label{fig:results_subline_images}
\end{figure*}

\begin{figure}
  \includesvg[width=1\columnwidth]{figures/improvements.svg}
  \caption{Fraction of improvement in total scan time over a scan with
    512 lines vs scan rate for the three sets of CS scans (at 25\%,
    15\% and 12.5\%) and courser raster scans at 128 and 64 lines.}
  \label{fig:improve_512}
\end{figure}

\section{Conclusions}\label{sec:conclusions}
The main conclusion of this paper is that with CS, it seems that we
can acquire images faster yet with less specimen damage
than is possible while raster scanning a full image.
On the other hand, it appears that reconstructing a course raster scan makes
more sense for faster scan rates. This is notable, since a course raster scan only
requires an additional step in post-processing. It remains to be seen if this conclusion
holds for a wide range of instruments.

%%%%%%%%% then the Bibliography, if any %%%%%%%%%
% \bibliographystyle{plain} % or "siam", or "alpha", etc.
\bibliographystyle{IEEEtran} % or "siam", or "alpha", etc.
% \nocite{*} % list all refs in database, cited or not
\bibliography{braker_bibs,Andersson_bibs,Yufans_bibs} % Bib database in "refs.bib"

%%%%%%%%% then the Appendices, if any %%%%%%%%%
\end{document}