### 2019-06-27 latex/coverletter.tex
Start coverletter

\documentclass[a4paper,twoside]{article}
\usepackage[margin=1in]{geometry}
\begin{document}
\hspace*{0.45\linewidth}
\begin{minipage}{0.52\linewidth}
Roger A. Braker\par
University of Colorado, Boulder\par
Dept of Electrical, Computer and Energy Engineering \par
425 UCB\par
Boulder, CO 80305\par
United States\par
roger.braker@colorado.edu\par
\today
\end{minipage}
\par\bigskip

Dear Prof. Qingze Zou,\par\bigskip
% check this is correct

It is our pleasure to submit this manuscript entitled ``Improving the Image Acquisition Rate of an Atomic Force Microscope Through Sub-sampling and Reconstruction'' (authored by Roger A. Braker Yufan Luo, Lucy Y. Pao and Sean Andersson) for consideration to be published in the Transactions on Mechatronics. This paper is submitted for possible publication in the Focused Section on Nano/Micro 
Motion  Systems:  Design,  Sensing  and  Control(NMMS).

\par\bigskip
\noindent Sincerely,
\par\bigskip
\noindent R.A. Braker\par
\noindent Y. Luo\par
\noindent L.Y. Pao
\noindent S. Andersson

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: