### 2019-10-07 latex/review-response-1.tex
Start rebuttal

\documentclass[11pt]{article} % use larger type; default would be 10pt
\usepackage{xr} % allow crossref to root.tex of all references
\externaldocument{jpaper}
\usepackage{xcite} % allow crossref to root.tex to all citations therein.
\externalcitedocument[A-]{jpaper}
%%% PACKAGES

% \usepackage[resetlabels,labeled]{multibib}
% \newcites{A}{review_response}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bbold}
\usepackage{mathtools}	
\usepackage{epstopdf}
\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{url}
\usepackage[margin=1in]{geometry}		%%DSZ: changed the margins 
\usepackage{appendix}
\usepackage{calc}
\usepackage{xspace}

\newcommand{\pic}[2]{\includegraphics[width={#2}]{#1}}
%% Try this out

\usepackage{enumitem}

\usepackage[normalem]{ulem}
\usepackage{soul}
\usepackage{framed}
\usepackage{environ}
\usepackage{xcolor}

\newcommand{\sgn}{\ensuremath{\text{sgn}}\xspace}
\newcommand{\sat}{\ensuremath{\text{sat}}\xspace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The following two macros allow us to use citations in this document which are
% not in the primary manuscript and add a prepended letter ('R' in this case)
% to disambugate from the citations refered to by xcite from the main manuscript.
% In order to use this must include a bibliography and style file at the end. 
\makeatletter
\renewcommand\@biblabel[1]{[R#1]}
\makeatother

\setlist{  
  listparindent=\parindent,
  parsep=0pt,
}
% \let\svbibcite\bibcite
% \def\bibcite#1#2{\svbibcite{#1}{#2}}
% \makeatletter
% \let\svbiblabel\@biblabel
% \def\@biblabel#1{\svbiblabel{R#1}}
% \makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newtheorem{prlm}{Problem}

\renewenvironment{leftbar}[1]
{%\begin{leftbar}%
    \def\FrameCommand
    {%
        {\color{#1}\vrule width 3pt}%
        \hspace{0pt}%must no space.
        \fboxsep=\FrameSep\colorbox{lightgray}%
    }%
    \MakeFramed{\hsize \hsize\advance\hsize-\width\FrameRestore}%
}
{\endMakeFramed}

% ---------------------------------------------------------------- %
% use this command as \listintertext{text here} to insert a rebuttal text
% between list items.
\makeatletter
\newcommand{\listintertext}{\@ifstar\listintertext@\listintertext@}

\newcommand{\listintertext@}[1]{% \listintertext*{#1}
  \hspace*{-\@totalleftmargin}
  % \begin{minipage}{\textwidth}
  \parbox{\textwidth}{
    \setlength{\parindent}{24pt}
    #1
    }}
  % \end{minipage}}

% \newcommand{\listintertext}{\@ifstar\listintertext@\listintertext@@}

% \newcommand{\listintertext@}[1]{% \listintertext*{#1}
%   \hspace*{-\@totalleftmargin}\parbox{\textwidth}{#1}}

% \newcommand{\listintertext@@}[1]{% \listintertext{#1}
%   \hspace{-\leftmargin}\mbox{#1}}
\makeatother

\newcommand{\inttext}[1]{\end{enumerate} #1\begin{enumerate}[resume]}
\newcommand{\itemaddr}[2]{\begin{leftbar}{#1} \item{#2}\end{leftbar}}

\title{Response to Reviews}
\author{Roger A. Braker and Lucy Y. Pao}

\date{\parbox{\linewidth}{\centering%
  %\today
\endgraf\bigskip
  }}

%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is prined

\begin{document}
\maketitle
\noindent
Dear Reviewers and Prof. Editors,

We thank the reviewers for their helpful comments. 
We have revised the paper based on these suggestions, with full details described below. The original review comments appear with a vertical bar on the left. References to equations use the numbers as they appear in the revised manuscript. 
%Citations specifically used in this rebuttal document are pre-pended with the letter 'R', e.g., [R1].

\section{Review 1}
\begin{leftbar}{black}
  A Comparison of 
\end{leftbar}

\begin{enumerate}

    \itemaddr{black}{Objection}

    \inttext{Response}
  \end{enumerate}
  While interesting, this paper assumes full state feedback.
  That is not a realistic assumption for many systems,
  including the one under consideration here.

\section{Review 2}

  \listintertext{Response. }

% \bibliographystyle{IEEEtran}
% \bibliography{review_response, prefixnumbers=A}

\bibliographystyle{IEEEtran}
% \bibliography{/home/arnold/bib_pdf/main_bibliography,review_response}
%\bibliography{afm_mpc_tcst.bib,review_response}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: